Liveness-Based Garbage Collection for Lazy Languages

### 2016-02-04 paper/paper.tex
Added reference to safe for space paper
references and space saving

\vspace*{-2mm}
\section{Related work} 
\label{sec:relatedwork}
The impact of  liveness on the effectiveness of  GC is
investigated in~\cite{Hirzel}.   The observation is that  liveness can
significantly  impact   garbage  collection,  but  only   when  it  is
interprocedural.  As far as memory requirement is concerned, our paper
demonstrates this observation.

There have been several attempts to use liveness analysis to improve 
GC for imperative languages.~\cite{khedker07heap}
presents a liveness analysis and uses the results for inserting 
nullifying statements in Java programs. In ~\cite{ran.shaham-sas03},
an automaton modelling heap accesses is created, and
temporal properties like liveness are checked  against the model.  
However both the approaches are restricted to intraprocedural scope.

In the  space of  functional languages,  there are:  rewriting methods
such                                                                as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based  analysis~\cite{tofte98region},  and insertion  of  compile-time
nullifying     statements~\cite{inoue88analysis,lee05static,Hamilton}.
All compile-time marking  approaches rely on an  efficient and precise
alias  analysis  and cannot  provide  significant  improvement in  its
absence.  An interesting approach suggested in~\cite{HofmannJ03} is to
annotate the heap usage of  first-order programs through linear types.
The  annotations  are  then  used to  serve  memory  requests  through
re-allocation.  However, this requires the user to write programs in a
specific way.

@@

%% \begin{figure}[t]
%% \centerline{\epsfig{file=lambda.eps, height=4cm, width=7cm}}
%%  \caption{Memory  usage and garbage collection pattern of  {\tt
%% lambda}.}
%% \label{fig:memory-usage-lambda} \figrule
%% \end{figure}
\vskip -5mm
\section{Future work and conclusions}
\label{sec:conclusion}
%% We have  defined a notion of  liveness on structured
%% data; this generalizes classical liveness and strong
%% liveness.    We  started   with   a  general   fully
%% context-sensitive analysis  which we  proved correct
%% with respect  to a minefield semantics  (this models
%% the  effect  of  garbage  collection  between  every
%% evaluation step).

>>>>

\vspace*{-2mm}
\section{Related work} 
\label{sec:relatedwork}
The impact of  liveness on the effectiveness of  GC is
investigated in~\cite{Hirzel}.   They observe that  liveness can
significantly  impact   garbage  collection,  but  only   when  it  is
interprocedural.  As far as memory requirement is concerned, our paper
demonstrates this observation.

There have been several attempts to use liveness analysis to improve 
GC for imperative languages.~\cite{khedker07heap}
presents a liveness analysis and uses the results for inserting 
nullifying statements in Java programs. In~\cite{ran.shaham-sas03},
temporal properties like liveness are checked  against an automaton modelling heap accesses.
Both these approaches are intraprocedural in scope.

In the  space of  functional languages,  there are:  rewriting methods
such                                                                as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based  analysis~\cite{tofte98region},  and insertion  of  compile-time
nullifying     statements~\cite{inoue88analysis,lee05static,Hamilton}.
All compile-time marking  approaches rely on an  efficient and precise
alias  analysis  and cannot  provide  significant  improvement in  its
absence.  An interesting approach suggested in~\cite{HofmannJ03} is to
annotate the heap usage of  first-order programs through linear types.
The  annotations  are  then  used to  serve  memory  requests  through
re-allocation.  However, this requires the user to write programs in a
specific way. 
%
Safe-for-space~\cite{appel.cps}              approaches~\cite{Clinger,
  shao00efficient}  reduce the  amount of  heap used  by a  program by
allocating   closures    in   registers   and   through    tail   call
optimizations. However, these approaches take care of only part of the
problem addressed  by our  analysis as the  program can  still contain
unused objects and closures that are reachable.
%

@@

%% \begin{figure}[t]
%% \centerline{\epsfig{file=lambda.eps, height=4cm, width=7cm}}
%%  \caption{Memory  usage and garbage collection pattern of  {\tt
%% lambda}.}
%% \label{fig:memory-usage-lambda} \figrule
%% \end{figure}
\vspace*{-2mm}
\vskip -5mm
\section{Future work and conclusions}
\label{sec:conclusion}
%% We have  defined a notion of  liveness on structured
%% data; this generalizes classical liveness and strong
%% liveness.    We  started   with   a  general   fully
%% context-sensitive analysis  which we  proved correct
%% with respect  to a minefield semantics  (this models
%% the  effect  of  garbage  collection  between  every
%% evaluation step).
